Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-25 VMCAI-revised-version/final.tex
effectivity

In 1986, Memmi and Vautherin introduced the (restricted) notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for $X$ a singleton. In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}). In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.

\noindent
{\bf Our contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
Moreover, {state-resilience},
{bounded-state-resilience} and
{$k$-state-resilience}
are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility. Moreover, the resilience problem is decidable for post-ideal-effective downward-WSTS with 
$\Safe$ upward-closed
%and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%%%
%
\item We clarify the different effectiness conditions that allow decidability of the six resilience.
%
\item We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property. We still extend  \cite{DBLP:journals/corr/PrasadZ16} as follows: the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
%
\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
%
%			$\mathop{\uparrow} \post^*(s)$ 
%
for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
%
\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
\end{itemize}

@@

@@

\section{Resilience for WSTS}

@@

\begin{proof}
\nocite{araki1976PN,dufourd1998reset}
{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
deciding structural termination of reset-VASS,
since termination is w.l.o.g. reachability of $q_f({\bf 0})$ 
 \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g. ??} 
for some control state $q_f$. \qed \end{proof}

@@

\alain{remercier les reviewers !} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.

>>>>

In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} in Petri nets for $X$ a singleton. In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}). In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.

\noindent
{\bf Our main contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X$ a singleton) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
%Moreover, {state-resilience},
%{bounded-state-resilience} and
%{$k$-state-resilience}
%are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
%$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item The three resilience problems are decidable for post-ideal-effective Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
%
\item  Resilience is also decidable, when 
$\Safe$ is upward-closed, for effective pred-basis WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
%
%    and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%				and for post-ideal-effective downward-WSTS ????
%%%
%
\item We clarify the different effectiness conditions on WBTS and WSTS that allow the decidability of the six resilience.
%
\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} ; we also show that relaxing some hypothesis leads to undecidability.
%
%			à utiliser dans le texte ???
%
%We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
%$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.

\item We show that the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities (Theorem \ref{downward srp} extends Theorem 1 in \cite{DBLP:journals/corr/PrasadZ16}).
The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WSTS with strong downward compatibility.
%
%			à utiliser dans le texte ???
%
%\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
%%
%%			$\mathop{\uparrow} \post^*(s)$ 
%%
%for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
%However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
%
\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
\end{itemize}

Most of our results about the different models and different types of resilience are in Figure YYYY.

@@

\alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}

@@

\alain{il manque des refs dans la fig 1 qui sont déjà dans la biblio: WBTS, omega2,on peut mettre les front lossy avec les LCM et LCS}
%
\section{Resilience for WSTS}

@@

\begin{proof}
\nocite{araki1976PN,dufourd1998reset}
{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
deciding structural termination of reset-VASS,
since termination is w.l.o.g. reachability of $q_f({\bf 0})$ 
 \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g. ??} 
\alain{qu'entends-tu par deterministic ici ?}
for some control state $q_f$. \qed \end{proof}

@@

\section*{Acknowledgements} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.
